---
title: "ERP plotting exercise"
output: html_notebook
---

In this exercise, you'll practice loading some files that have already been processed, creating summary statistics, and running simple statistical tests.

You'll need the following packages: 
* tidyverse

We'll be using the following commands
* read_csv (or "Import Dataset")
* gather()
* %>%
* separate()

```{r}
library(tidyverse)
```

Import the data from the pre-prepared .csv.

```{r}
ObjFreqERPs <- read_csv("../data/ObjFreqERPs.csv")
```

```{r}
glimpse(ObjFreqERPs)
```

The data is currently in long format. Use gather() to create two new columns - "condition" and "amplitude", keeping the columns "Time" and "Subject".

```{r}
ObjFreqERPs %>%
  gather(condition, amplitude, -Subject, -Time)
```

The data is now in long, tidy format: each column is one variable, each row is one observation. However, we need to split "condition" into two variables, one for each factor. Remember that this is a 2 x 2 design. Use the separate() command.

```{r}
final_df <- ObjFreqERPs %>%
  gather(condition, amplitude, -Subject, -Time) %>%
  separate(condition, into = c("Object", "Frequency"), sep = "/")
head(final_df)
```

Save the new dataframe.

```{r}
save(final_df, file = '../user_data/final_df.RData')
```
