---
title: "How to run standard statistics"
output: html_notebook
---

In the first exercise, you practiced plotting ERPs and many data manipulation commands. 

Now we'll move on to running some straightforward statistics.

Remember, help for any command can be looked up using ?command in console, or searching for it in the help pane.

## Packages and commands

You'll need to load the following packages
* tidyverse
* afex

We'll be using the following new commands
* group_by()
* summarise()
* t.test()
* aov_ez()
* map()/map_dbl()

```{r}
library(tidyverse)
library(afex)
```

## Pre-specified time windows

It's common to choose on a priori grounds specific time windows for analysis, taking, for example, mean amplitudes over a particular time window. 

Let's start by examining ERPs in the P1 time window. We'll begin by pretending this isn't a 2 X 3 design, and doing paired t-tests on mean amplitudes in the time window from approximately 85-135 ms. 



```{r}
final_df %>%
  filter(Time >= 85 & Time <= 135) %>%
  group_by(Subject,Object) %>%
  summarise(amplitude = mean(amplitude))
```