---
title: "R basics"
output: html_notebook
---





## Assignment

The assignment operator in R is <-

```
a <- 2 + 2
```

Note = 

## Functions

```
a <- mean(x)

```

## Installing and loading packages

Packages extend the capabilites of R by implementing functions. Functions do things such as calculating means, run ANOVA.

Several of the packages we'll be using have already been installed.

Packages which are installed are loaded using the *library()* command.

Use *library()* to load the **tidyverse** package.

```{r load_packages}
library(tidyverse)
```

The **tidyverse** package is actually a collection of several packages that implement a wide variety of features that we'll be using.

One of those is **dplyr**, a package for manipulation and reorganization of data.

Packages often include help or information in the form of *vignettes*; short documents giving more detail about the purpose of and use of the package.

Those available with a package can be found using *browseVignettes()*. The package name must be given in quotes.

Try this now for **dplyr**.

```{r vignette}
browseVignettes("dplyr")
```

**dplyr** includes functions such as *filter()*, *mutate()*, and *select()*.

Help for functions can be found using *?function_name* if the package is already loaded, *??function_name* if it is not.


```{r help_function}
?mutate
```
