---
title: "Fitting and plotting individual GLMs"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
---

One way of modelling stingle trial data is to fit individual GLMs.

## Packages and commands

packages:
* feather
* tidyverse

New commands:
* read_feather()

## Loading Feather files

First, let's load some data. We'll be loading some data from a *feather* file using the **feather** package. *feather* is a lightweight binary format; the data has been pre-processed using MNE-Python and exported from there.

Choose one of the feather files from the data directory, and load it using *read_feather()*

```{r load_packages, message = FALSE}
library(feather)
library(tidyverse)
feather_dat <- read_feather('../data/S104Extero.feather')
head(feather_dat)
```

Rearrange the data  to long format. Trim it down to the time range -100 to 400 ms using *filter()*, use *separate()* to split the condition column into multiple columns. Use *gather()* bring the individual electrode columns into a key-value pair.

```{r reorganize_data}
feather_dat <- feather_dat %>%
  filter(time >= -100 & time <= 400) %>%
  separate(condition, into = c("light", "touch", "reported"), sep = "/") %>% 
  gather(electrode, amplitude, -time, -touch,-light,-reported, -epoch)
head(feather_dat)
```

## Plot ERPS

Now plot the ERPs from all electrodes and conditions.

```{r plot_ERPs}
ggplot(feather_dat, aes( x = time, y= amplitude, colour = electrode)) +
  stat_summary(geom="line") +
  theme_classic()
```

## Nest and run GLM

Nest the data and calculate a linear model on every timepoint at every electrode.

```{r run_glms}
feather_dat <- nest(feather_dat, -time, -electrode)
feather_dat <- feather_dat %>%
  mutate(fit = map(data, ~lm(amplitude ~ touch*light, data = .)))

head(feather_dat)
```

## Plot R-Squared

We can extract the summary of the model, and the R-Squared from that model.

```{r get_r2}
feather_dat <- feather_dat %>%
  mutate(summ = map(fit,~summary(.)),
         r2 = map_dbl(summ,"r.squared"))

ggplot(feather_dat, aes(x = time, y = r2, colour = electrode)) +
  geom_line() +
  theme_classic()

```


