---
title: "Running Linear mixed-effects models"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
---

As a final step, we'll load multiple files and run a linear mixed effects model 

## Packages and commands

Load the following packages:
* tidyverse
* lmerTest
* feather

New commands we'll be using are:
* list.files()
* lmer()
* map_df()

```{r load_packages, message= FALSE}
library(feather)
library(tidyverse)
library(lmerTest)
library(eegUtils)
library(sjPlot)
options(contrasts = c("contr.sum","contr.poly"))
```

## Loading multiple files

In **../data** folder you'll find a number of *.feather* files. Since these all have the same structure and are the same filetype, we can load them all that once and combine them into a single data frame using *map_df()*. *list.files()* can be used to generate a list of files which meet a particular pattern within a folder. Here we'll load all the *.feather files.


```{r load_all, warning = FALSE}
all_subs <- map_df(list.files('../data/','*.feather', full.names = TRUE),
                   ~read_feather(.x),
                   .id = "Subject")
head(all_subs)
```

Let's make the data a little more manageable and convert it to long format. Use *filter()* to choose data from -100 to 400 ms; *separate()* to split up the condition column; and *gather()* to convert the data to long format fully. 

```{r long_format}
all_subs <- all_subs %>%
  filter(time >= -100 & time <= 400) %>%
  separate(condition, into = c("light", "touch", "reported"), sep = "/") %>%
  gather(electrode, amplitude, Fp1:O2)

head(all_subs)
```

## Summarise and plot grand mean ERPs

It is extremely simple to produce grand means; we'll do this in a single chain of piped commands, since we do not want to keep teh grand averaged data - for LMEs, we need single trials.

```{r grand_means}
all_subs %>% 
  group_by(Subject, electrode, time) %>%
  summarise(amplitude = mean(amplitude)) %>%
  ggplot(aes(x = time, y = amplitude, colour = electrode)) +
  stat_summary(geom = "line") +
  theme_classic()
```

## Baseline correction

The data here needs to be baseline corrected. We'll calculate the baseline for each subject and electrode and use *left_join()* to add it to the original data frame.

```{r baseline}
all_subs <- left_join(all_subs, all_subs %>%
  filter(time >= -100 & time <= 0) %>%
  group_by(Subject, electrode) %>% 
  summarise(baseline = mean(amplitude)),
  by = c("Subject","electrode")) 

head(all_subs)
all_subs$amplitude <- all_subs$amplitude - all_subs$baseline
```

Now we'll plot the basline corrected data.

```{r plot_baselined}
all_subs %>% 
  group_by(Subject, electrode, time) %>%
  summarise(amplitude = mean(amplitude)) %>%
  ggplot(aes(x = time, y = amplitude, colour = electrode)) +
  stat_summary(geom = "line") +
  theme_classic()
```

## Select data for LME

LMEs can take quite a while to run (amongst other things...). For our workshop purposes, let's choose a specific time window to average over and a single electrode of interest. We'll pick 140 to 170 ms.

```{r cut_down}
mini_data <- all_subs %>%
  filter(time >= 140 & time <= 170) %>%
  group_by(Subject,touch,light,epoch, electrode) %>%
  summarise(amplitude = mean(amplitude)) %>%
  ungroup()
```

Let's use *topoplot()* from **eegUtils**. 

```{r topoplot, message = FALSE}
mini_data %>%
  group_by(Subject,electrode) %>%
  summarise(amplitude = mean(amplitude)) %>%
  topoplot()
```

Looks like we have something going on at occipital electrodes. Let's focus in on one electrode, PO4.

## Run an LME

```{r run_lme}
lme_out <- mini_data %>%
  filter(electrode == "PO4") %>%
  group_by(Subject, light, touch, epoch) %>%
  summarise(amplitude = mean(amplitude)) %>%
  ungroup() %>%
  lmer(amplitude ~ touch*light + (1 + touch*light|Subject), data = .)
```

We can use summary to have a look at the results.

```{r lme_output}
summary(lme_out)
```

## Plot fixed and random effects

```{r sjplots, message = FALSE}
sjp.lmer(lme_out, type = "fe", p.kr = FALSE)
sjp.lmer(lme_out, type = "re", sort.est = "(Intercept)")
```

We can try to run a full LME on every timepoint. Let's see what happens.

```{r warnings}
mo_problems <- all_subs %>%
  filter(electrode == "PO4") %>%
  ungroup() %>%
  nest(-time) %>%
  mutate(lme_out = map(data, ~lmer(amplitude~touch*light + (1 + touch*light |Subject), data = .)))
```

You will note a lot of error messages. 

